---
title: "Military Education Exchange - Network Analysis"
author: "Erin Lemons, Kenton Thibaut, Bria Pullin"
date: "Nov 30, 2021"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
editor_options: 
  chunk_output_type: console
---

<style> body {text-align: justify} </style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "",  echo = TRUE, message = F, warning = F)
library(knitr)
library(igraph)
library(visNetwork)
```

## Read Data
```{r}
#read in Matrix
f <- read.csv("adjacencyMatrixFormat.csv",header = T)
f <- cbind(f[,48],f) #paste last column to front
f <- f[,-49] #remove the last column

#Adjacency Matrix construction
nodelist <- names(f)[-1] 
my_matrix <- as.matrix(f) [,-1] 
rownames(my_matrix) <- colnames(my_matrix) <- nodelist 
my_matrix[1:5,1:5] 
```

The adjacency matrix shows the connections between different countries. A connection is recorded if a country's military leader has engaged in military education in a specified country. A non-zero value means a leader from the country in a particular row  has engaged in military education exchange with the country in the columns, while a value of zero indicates the opposite (a zero is also recorded for same-country pairs - e.g. USA and USA - because there are no self-loops allowed in this network). 

## Network Construction
```{r}
g <- graph.adjacency(my_matrix, mode = "Undirected", weighted = T)
g <- simplify(g,remove.multiple = T,remove.loops = T)
g
```

As we see in the above matrix, the relationship is symmetric - i.e., there are directed edges between two countries. The graph is also undirected, which affects the assigning of the weights. For example, the connection between USA--Russia has weight 2, and between USA--France has weight 6; Similarly, Russia--USA also has weight 2 and France--USA also has weight 6. The strength of ties between the different countries have different numeric values ranging from 0 to 9, with 9 being the strongest tie. 

In the graph, we can see the set of edges, and the number of nodes and edges. Here, nodes correspond to countries and edges correspond to military education exchange between countries (i.e., whether a leader from a particular country has engaged military education exchange with another country). 

## Network Statistics

- The *density* of a network is the ratio of the number of edges to the the number of possible edges. In other words, the proportion of present edges to all possible edges in the network. A dense network indicates a high number of ties between many of the nodes. A less dense network means that there are fewer shared ties among all the nodes. 

- *Transitivity* is a measure of the tendency of the nodes to cluster together. High transitivity means that the network contains communities or groups of nodes that are densely connected internally.

- *Network diameter* is the longest geodesic distance (length of the shortest path between two nodes) in the network. In igraph, diameter() returns the distance, while get_diameter() returns the nodes along the first found path of that distance. Note that edge weights are used by default, unless set to NA.

- *Average path length* is the mean of the shortest distance between each pair of nodes in the network (in both directions for directed graphs).

```{r}
#Total nodes: 47
vcount(g)

#Total edges: 156
ecount(g)

#Density of graph: 0.1443
graph.density(g,loops = F)

#Transitivity: 0.415
transitivity(g,type = "global")
triad.census(g)

#Diameter: weighted is 5, unweighted is 4
diameter(g)
diameter(g, weights = NA)
diam <- get_diameter(g)
diam 

vcol <- rep("gray40", vcount(g))
vcol[diam] <- "gold"
ecol <- rep("gray80", ecount(g))
ecol[E(g, path=diam)] <- "orange"
plot(g,layout=layout.sphere, vertex.color=vcol, edge.color=ecol, edge.arrow.mode=0, vertex.label=NA, main = "Military Education Network: Diameter\n(Longest Paths in Yellow)")

#degree distribution
deg_dist <- degree_distribution(g, cumulative=T, mode="all")
plot(x=0:max(degree(g)), y=1-deg_dist, pch=19, cex=1.2, col="orange",
xlab="Degree", ylab="Cumulative Frequency", main = "Military Education Network: Degree Distribution")

#mean distance
mean_distance(g)
distances(g) # with edge weights
distances(g, weights=NA) # ignore weights

#cliques
vcol <- rep("grey80", vcount(g))
vcol[unlist(largest_cliques(g))] <- "gold"
plot(as.undirected(g), vertex.label=NA, vertex.color=vcol, main = "Military Education Network: Cliques")

## Connected Components
components(g,mode = "strong") #Strongly Connected components

##
```

- According to the above network statistics, we know that the network contains 47 nodes (countries) & 156 edges (leader has engaged in military education exchange with country i).

- The density of the military education network is 0.14, which is low. This indicates that the network is made up of only a few nodes that are connecting to most of the network, with a larger number of nodes that are poorly connected. This low density is a result of the fact that many of the countries in the dataset have no connections in terms of military education exchange. 

- The transitivity of the network is 0.41. Perfect transitivity implies that, if $x$ is connected (through an edge) to $y$, and $y$ is connected to $z$, then $x$ is connected to $z$ as well. However, partial transitivity is useful. In many networks, particularly social ones, the fact that $x$ knows $y$ and $y$ knows $z$ does not guarantee that $x$ knows $z$ as well, but makes it much more likely. In this graph, the transitivity score indicates that 41% of the nodes are transitively connected to each other. 

- The diameter of the network is 5, which means the maximum shortest path in the network has 5 nodes in it. In other words, the maximum distance between any pair of nodes is 5. To find the diameter of the graph, we first find the shortest path between each pair of vertices. The greatest length of any of these paths is the diameter of the graph. The countries included in the diameter are `FME_Zimbabwe, FME_UK, FME_South_Africa, FME_Israel, FME_Senegal` and `FME_Morocco`. The diameter value shows that most of the developed countries are closer to each other, meaning they engage in  military education exchange more frequently with each other.

- The plot of the degree distribution of the network has the Y-axis representing the Frequency and the X-axis representing the Degree. In the study of graphs and networks, the degree of a node in a network is the number of connections it has to other nodes and the degree distribution is the probability distribution of these degrees over the whole network. The plot of the military network indicates that the network is composed of a few nodes with high degrees while most of the nodes have low degrees. This indicates that a small number of the countries - namely, more developed ones - in the network are supplying the majority of military education exchanges.. 

- In network analysis, a clique is a sub-set of a network in which the actors are more closely and intensely tied to one another than they are to other members of the network.The plot here shows the largest clique in the military education exchange network, which shows the clique with the the most concentrated degree distribution & density. From the clique network we see that either countries are exchanging military education with those countries who have a high number of connections, or with those countries who are close to it. 

- Futher, we see that the graph has 3 connected components, in which the largest component has size 45, and the other two components are comprised of single isolates.Given our node size is 47, this means that 2 nodes in the network have no connections at all, while the rest of the 45 nodes have connections in the network. 

## Network Visualization
```{r}
wk <- walktrap.community(g)
df <- as.data.frame(table(wk$membership))
names(df) <- c("Community Number", "Size")
kable(df,caption = "Communities in the Network")

nodes <- data.frame(id = V(g)$name, title = V(g)$name, group = wk$membership)
edges <- get.data.frame(g, what="edges")[1:2]

vis.nodes <- nodes
vis.links <- edges
vis.nodes$shape  <- "dot"  
vis.nodes$shadow <- TRUE 
vis.nodes$title  <- V(g)$name 
vis.nodes$label  <- V(g)$name 
vis.nodes$size   <- 35 
vis.nodes$borderWidth <- 2 

vis.links$width <- 3 
vis.links$font.size <- 15
vis.links$label <- E(g)$weight
vis.links$dashes <- as.vector(ifelse(E(g)$weight==1,TRUE,FALSE))
vis.links$color <- palette()[E(g)$weight]
vis.links$arrows <- "to" # arrows: 'from', 'to', or 'middle'
vis.links$smooth <- FALSE    
vis.links$shadow <- FALSE   
vis.nodes$color.border <- "black"
vis.nodes$color.highlight.background <- "orange"
vis.nodes$color.highlight.border <- "darkred"

 k <- visNetwork(vis.nodes, vis.links, width = "100%", height = 700) %>%
visIgraphLayout(layout = "layout.sphere")
 
ledges <- data.frame(color = palette()[unique(E(g)$weight)], label =unique(E(g)$weight))  
visLegend(k, main="Communities",addEdges = ledges, position="left", ncol=3)

plot(wk,g, vertex.label="", main = "Military Education Exchange Network: Communities") #walk trap community algorithm
```

The above network visualization shows the different communities in the military education exchange network. Qualitatively, a community is defined as a subset of nodes within the graph such that connections between the nodes are denser than connections with the rest of the network. We can see from the visualization & table that 3 or 4 communities have a large number of nodes, while the rest of the communities have only 1 or 2 nodes. The largest community in the network is presented in light red & the second largest is presented in light blue. From the visualization, it seems that most of the countries in the network tend to exchange military education with countries within their "circle." Looking into the larger  communities, we can see that the nodes within them have many connections within the community and very few connections outside the community. The top two countries, the USA and China, are in the same community. The "zoomed out" visualization of the communities in the network shows more clearly that nodes tend to form ties within their own communities; i.e., countries engaged in military education exchanges with countries within its circle.

## Centrality

In graph theory and network analysis, methods of centrality assign numbers or rankings to nodes within a graph corresponding to their network position. Applications include identifying the most influential person(s) in a social network, key infrastructure nodes in the Internet or urban networks, super-spreaders of disease, and brain networks, for example. Centrality measures are explored below as a way to identify the most influential countries in our network.

```{r centrality}
#Centrality table
Degree <- degree(g,mode = "all")
Betweenness <- betweenness(g)
Closeness <- closeness(g)
df <- data.frame(Degree, Betweenness, Closeness)
kable(df,caption = "Centrality Score for each Network")
```

Centrality measures are computed for each node, presented in the table. Here, we compute `Degree`, `Betweenness` and `Closenness` centrality (explained further below). The results show that the most important nodes in the network are `FME_USA` and `FME_China`. Each centrality measure is depicted in a network plot, where the size of the node represents its centrality score. The top 5 most important countries with regards to centrality measures are depicted below.

### Degree Centrality

The degree centrality of a node is simply its degree—the number of edges it has. The higher the degree, the more central the node is. This can be an effective measure, since many nodes with high degrees also have high centrality by other measures.

In our network, the top five nodes by degree centrality are the US, China, France, the UK, and Russia, in that order.

In the visualiztion below, node size is correlated with degree centrality. We can see that `USA` and `China` have the highest high number of military education exchanges with other countries.  


```{r}
#Top 5 nodes by Degree centrality
sort(degree(g,mode = "all"),decreasing = TRUE)[1:5]


## Network Visualization By Degree
nodes <- data.frame(id = V(g)$name, title = V(g)$name, group = wk$membership)
edges <- get.data.frame(g, what="edges")[1:2]

vis.nodes <- nodes
vis.links <- edges
vis.nodes$shape  <- "dot"  
vis.nodes$shadow <- TRUE 
vis.nodes$title  <- V(g)$name 
vis.nodes$label  <- V(g)$name 
vis.nodes$size   <- Degree * 3 
vis.nodes$borderWidth <- 2 

vis.links$width <- 3 
vis.links$font.size <- 15
vis.links$label <- E(g)$weight
vis.links$dashes <- as.vector(ifelse(E(g)$weight==1,TRUE,FALSE))
vis.links$color <- palette()[E(g)$weight]
vis.links$arrows <- "to" 
vis.links$smooth <- FALSE    
vis.links$shadow <- FALSE   
vis.nodes$color.border <- "black"
vis.nodes$color.highlight.background <- "orange"
vis.nodes$color.highlight.border <- "darkred"

 k <- visNetwork(vis.nodes, vis.links, width = "100%", height = 700, main = "Network Visualization by Degree centrality") %>%
visIgraphLayout(layout = "layout.sphere")
 
ledges <- data.frame(color = palette()[unique(E(g)$weight)], label =unique(E(g)$weight))  
visLegend(k, main="Communities",addEdges = ledges, position="left", ncol=3)

```


### Betweenness Centrality
Betweenness centrality quantifies the number of times a node acts as a bridge along the shortest path between two other nodes, that is, betweenness centrality ranks the importance of a node in terms of its role as a "bridge"connecting other nodes.

The top five nodes by betweenness centrality in this network are the US, China, Tanzania, Angola, and South Africa, in that order. 

In the visualization below, we can see that the `USA` and `China` are the top ranked in terms of betweenness centrality. In other words, countries tend to engage in military education exchange with countries that are connected (i.e, have engaged in education exchange) with the US and/or China. 


```{r}
#Top 5 nodes by Betweenness Centrality
sort(betweenness(g,directed = FALSE),decreasing = TRUE)[1:5]


## Network Visualization By Betweenness
nodes <- data.frame(id = V(g)$name, title = V(g)$name, group = wk$membership)
edges <- get.data.frame(g, what="edges")[1:2]

vis.nodes <- nodes
vis.links <- edges
vis.nodes$shape  <- "dot"  
vis.nodes$shadow <- TRUE #Drop node shadow
vis.nodes$title  <- V(g)$name # Text on click
vis.nodes$label  <- V(g)$name # Node label
vis.nodes$size   <- Betweenness * 0.8 # Node size
vis.nodes$borderWidth <- 2 # Node border width

vis.links$width <- 3 # line width
vis.links$font.size <- 15
vis.links$label <- E(g)$weight
vis.links$dashes <- as.vector(ifelse(E(g)$weight==1,TRUE,FALSE))
vis.links$color <- palette()[E(g)$weight]
vis.links$arrows <- "to" # arrows: 'from', 'to', or 'middle'
vis.links$smooth <- FALSE    #edges curved? Y/N
vis.links$shadow <- FALSE    # edge shadow? Y/N
vis.nodes$color.border <- "black"
vis.nodes$color.highlight.background <- "orange"
vis.nodes$color.highlight.border <- "darkred"

 k <- visNetwork(vis.nodes, vis.links, width = "100%", height = 700, main = "Network Visualization by Betweenness centrality") %>%
visIgraphLayout(layout = "layout.sphere")
 
ledges <- data.frame(color = palette()[unique(E(g)$weight)], label =unique(E(g)$weight))  
visLegend(k, main="Communities",addEdges = ledges, position="left", ncol=3)
```


### Closeness Centrality

Closeness centrality indicates how close a node is to all other nodes in the network. It is calculated as the average of the shortest path length from the node to every other node in the network.

In our network, the top five nodes by closeness centrality are China, the US, South Africa, Angola, and Brazil, in that order. 



```{r}
#Top 5 nodes by closeness centrality
sort(closeness(g,mode = "all"),decreasing = TRUE)[1:5]

## Network Visualization By Closeness
nodes <- data.frame(id = V(g)$name, title = V(g)$name, group = wk$membership)
edges <- get.data.frame(g, what="edges")[1:2]

vis.nodes <- nodes
vis.links <- edges
vis.nodes$shape  <- "dot"  
vis.nodes$shadow <- TRUE 
vis.nodes$title  <- V(g)$name 
vis.nodes$label  <- V(g)$name 
vis.nodes$size   <- Closeness * 10000 
vis.nodes$borderWidth <- 2 

vis.links$width <- 3 
vis.links$font.size <- 15
vis.links$label <- E(g)$weight
vis.links$dashes <- as.vector(ifelse(E(g)$weight==1,TRUE,FALSE))
vis.links$color <- palette()[E(g)$weight]
vis.links$arrows <- "to" 
vis.links$smooth <- FALSE    
vis.links$shadow <- FALSE
vis.nodes$color.border <- "black"
vis.nodes$color.highlight.background <- "orange"
vis.nodes$color.highlight.border <- "darkred"

 k <- visNetwork(vis.nodes, vis.links, width = "100%", height = 700, main = "Network Visualization by Closeness Centrality") %>%
visIgraphLayout(layout = "layout.sphere")
 
ledges <- data.frame(color = palette()[unique(E(g)$weight)], label =unique(E(g)$weight))  
visLegend(k, main="Communities",addEdges = ledges, position="left", ncol=3)
```



































